---
title: "MetaVolcano_metaanalysis"
output: html_notebook
---

##Background
Meta analysis of the dataset according to the MetaVolcanoR package. Intitial analysis performed in June 2021 by Beat Moeckli and edited in November 2021.

First import of previously created df_result file with all result in a single dataframe and preparation for use with the MetaVolcano package

```{r import of dataset and preparation}
library(MetaVolcanoR)
library(dplyr)

#Import dataframe
df_results<-read.csv("df_results.csv")%>%
  dplyr::filter(!GEOSET%in%c("GSE133767", "GSE62715")) #Filter datasets from fetal and blastocyst offspring
list_results<-split(df_results, f=df_results$GEOSET)

```

## Create the metaVolcano elements for further analysis, first random effect model

Random effect model analysis, The summary p-value stands for for the probability that the summary fold change is not different than Zero the perturbation ranking follows the topconfect approach

```{r Random effect model according to top confect}
meta_degs_rem <- rem_mv(diffexp=list_results,
                        pcriteria="P.Value",
                        foldchangecol='logFC', 
                        genenamecol='symbol',
                        geneidcol=NULL,
                        collaps=TRUE,
                        llcol='CI.L',
                        rlcol='CI.R',
                        vcol=NULL, 
                        cvar=TRUE,
                        metathr=0.01,
                        jobname="MetaVolcano",
                        outputfolder=".", 
                        draw='HTML',
                        ncores=1)

#Gives the top perturbated genes, following the topconfect approach
head(meta_degs_rem@metaresult, 40)

#Draws MetaVolcano Plot
meta_degs_rem@MetaVolcano

#Draws forest plot for individual genes with with all the studies
draw_forest(remres=meta_degs_rem,
            gene="Fgf21",
            genecol="symbol", 
            foldchangecol="logFC",
            llcol="CI.L", 
            rlcol="CI.R",
            jobname="MetaVolcano",
            outputfolder=".",
            draw="HTML")
```

## Vote Count model 
Assesses in how many of the datasets genes are differentially expressed. Criteria for the genes to be differentially expressed can be manually set.

```{r Vote count model}
meta_degs_vote <- votecount_mv(diffexp=list_results,
                               pcriteria="P.Value",
                               foldchangecol='logFC', 
                               genenamecol='symbol',
                               collaps=T,
                               pvalue=0.05,
                               foldchange=0, 
                               metathr=0.01)

#Gives the frequency of differentially regulated genes in each dataset
meta_degs_vote@degfreq
head(meta_degs_vote@metaresult, 1000)

#Vulcano Graph showing consistency of differential expression
meta_degs_vote@MetaVolcano

```

## Combining-approach
Analysis combining fold change and p-value combination. "idx" is calculated as metafc*-log10(metap), this according to the topconfect method

```{r combining approach}
meta_degs_comb <- combining_mv(diffexp=list_results,
                               pcriteria='P.Value', 
                               foldchangecol='logFC',
                               genenamecol='symbol',
                               geneidcol=NULL,
                               metafc='Mean',
                               metathr=0.01, 
                               collaps=TRUE,
                               jobname="MetaVolcano",
                               outputfolder=".",
                               draw='HTML')

head(meta_degs_comb@metaresult, 500)
meta_degs_comb@MetaVolcano
```


